---
title: "Data Wrangling"
author: "Faizal Ramli"
date: "6/6/2020"
output:
  md_document:
    variant: markdown_github
---

## Working with tidyr Package
In this first section I tried to use `gather()` and `spread()` function from `tidyr` package in order to convert the structure of a data frame to be more tidy. I use three data frames from `EDAWR` package which are `storm`, `cases`, and `pollution` .
```{r}
library(EDAWR)
storms
storms$storm
storms$wind
storms$pressure
storms$date

cases
cases$country
names(cases)[-1]
unlist(cases[1:3, 2:4])

pollution
pollution$city[c(1, 3, 5)]
pollution$amount[c(1, 3, 5)]
pollution$amount[c(2, 4, 6)]

library(tidyr)
gather(cases, "year", "n", 2:4)
gather(cases, "year", "n", 2:4) %>% spread(year, n)

spread(pollution, size, amount)
spread(pollution, size, amount) %>% gather("size", "amount", 2:3)
```
Another functions in `tidyr` package that would be useful are `separate()` and `unite()` to modify the structure of value on a varibel. Below are two examples on how to use these functions.
```{r}
separate(storms, date, c("year", "month", "day"), sep = "-")
separate(storms, date, c("year", "month", "day"), sep = "-") %>%
  unite("date", day, month, year, sep = "/")
```

## Working with dplyr Package
`dplyr` is a powerful package for doing data wrangling in R, it contains many functions that are useful for R user to manipulate the dataset such are `select()` for selecting any variabel and `filter()` for filtering any value in some conditions, and many others. Examples on how to use these functions can be seen below.
```{r}
library(dplyr)
EDAWR::storms
select(EDAWR::storms, storm, pressure)
select(EDAWR::storms, -storm)
select(EDAWR::storms, wind:date)
EDAWR::storms %>% select(contains("ate"))
EDAWR::storms %>% select(ends_with("e"))
EDAWR::storms %>% select(starts_with("s"))

EDAWR::storms %>% filter(pressure > 1005, wind >= 45)
EDAWR::storms %>% filter(pressure > 1005, wind >= 45,
                         storm %in% c("Alberto", "Alex"))
EDAWR::storms %>%
  filter(pressure > 1005, wind >= 45, storm %in% c("Alberto", "Alex")) %>%
  mutate(ratio = pressure/wind)
EDAWR::storms %>%
  filter(pressure > 1005, wind >= 45, storm %in% c("Alberto", "Alex")) %>%
  mutate(ratio = pressure/wind, inverse = ratio^-1)

pollution %>% summarise(median = median(amount), variance = var(amount))
EDAWR::storms %>% arrange(wind)
EDAWR::storms %>% arrange(desc(wind))

pollution %>% group_by(city) %>%
  summarise(mean = mean(amount), sum = sum(amount), n = n())
pollution %>% group_by(size) %>%
  summarise(mean = mean(amount), sum = sum(amount), n = n())

tb
tb %>% group_by(country, sex) %>% summarise(n = n())
```
With `dplyr` package we can also join two table with many functions such are `right_join()`, `left_join()`, `full_join()`, and many others. Here are some examples on how to join tables with `dplyr`.
```{r}
y
z
bind_cols(y, z)
bind_rows(y, z)
union(y, z)
intersect(y, z)
setdiff(y, z)

songs
artists
left_join(songs, artists, by = "name")
right_join(songs, artists, by = "name")
inner_join(songs, artists, by = "name")
full_join(songs, artists, by = "name")
semi_join(songs, artists, by = "name")
semi_join(artists, songs, by = "name")
anti_join(songs, artists, by = "name")
anti_join(artists, songs, by = "name")

songs2
artists2
left_join(songs2, artists2, by = c("first", "last"))
```